\bf Excuse me, do you have a moment to talk about version control?

### 2017-07-07 README.Rmd
Abstract, conclusion, Header 1 not 2, spellcheck

---
title: "Excuse me, do you have a moment to talk about version control?"
authors:
- name: Jennifer Bryan
  affiliation: |
    RStudio and the Department of Statistics, University of British Columbia
  thanks: |
    The author gratefully acknowledges the constructive feedback from reviewers
    Nicholas Horton, Colin Rundel, and Hadley Wickham.
blinded: 0
output:
  github_document:
    toc: yes
  rticles::asa_article: null
keywords: version control
abstract: Abstract abstract abstract.
bibliography: git-github-for-stats.bib
---

```{r knitr_options, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, out.width = "100%")
```

```{r include = FALSE}
## I temporarily embed some figures from the internet and LaTeX just. can't.
## deal. with. it.
is_github_document <-
  knitr::opts_knit$get("rmarkdown.pandoc.to") == "markdown_github"
```

## Why Git?

@@

## What is Git?

@@

## Who should read this and what to expect

@@

## What is GitHub?

@@

## Initial system setup

@@

  * I repeat: consider using a graphical front-end for Git, a.k.a. a Git client, versus restricting yourself to the command line interface.
  * Establish confidence in the basics (e.g. make a change, commit it, push it) before wading into more advanced usage (e.g. branching).
  * Commit yourself to Git usage on a project that will provide sustained practice over several months. Usage in a course is great, because it provides a relentless stream of small deadlines.
  * Realize that no one is giving out Git style points. It's ok to "power-cycle", i.e. re-initialize the Git repository, to get unstuck.

## Repositories and workflow

@@

## Commits, diffs, and tags

@@

## Markdown is special on GitHub

@@

## Markdown is special for R users

@@

Before we move on, I want to zoom out and revisit R Markdown and the rmarkdown package more generally. Note that R Markdown can be rendered to many more formats than Markdown, including HTML, PDF, and Microsoft Word, and can incorporate code chunks in many languages other than R. Markdown is emphasized here because it is extraordinarily useful in GitHub-hosted projects. Many people are used to targetting other output formats and overlook this special synergy between Markdown and GitHub.

## Which files to commit

@@

__Intermediates:__ These files are programmatically generated and serve a temporary purpose, but are not intrinsically valuable (e.g., `.aux` and `.log` in LaTex workflows).

@@

## Collaboration

@@

## GitHub as web presence

@@

## More resources

I've tried to convey the main points about the use of Git and GitHub in statistical and data analytical settings, but I've had to leave many things out. There are more advanced topics that will come up once your use of Git becomes more sophisticated and there are topics that are only relevant to certain types of reader.

@@

Branches and pull requests are an extremely powerful feature of Git/GitHub and should be your first foray beyond the basics of commit, push, and pull. A branch is an independent line of development within a repo, where the intent is usually to merge it into the master branch when ready. In a course website, you might work in a branch to update a series of lessons, while leaving the current version intact in the meantime. A pull request is a GitHub-specific way to propose changes to a repo that augments Git's regular branch and merge workflow with facilities for structured review. A pull request can be made between two branches in the same repo, such as `master` and `bugfix`, or between two independent copies of a repo, an original and a so-called "fork". This is the mechanism for making and accepting contributions to open source software projects on GitHub, including many popular R packages *link to a search?*.

*inspiration for possible figure on collaboration/mergin, from the barlett talk*

```{r echo = FALSE, out.width="100%"}
knitr::include_graphics("img/bartlett-merge-commit.png")
```

## Conclusion

*NEEDS ONE ... BUT SHORT ... SOME FODDER*

Transparency about process and product is increasingly important in science. The SOMETHING for reproducibility is well accepted. A more underappreciated benefit is democratization of our field, as this affords a much broader audience a clear view of how scientists and programmers work.

make statistical thought and implementation available

```{r eval = is_github_document, echo = FALSE, results = 'asis'}
## surely there is a better way?
cat("## References\n")
```

>>>>

---
title: "Excuse me, do you have a moment to talk about version control?"
authors:
- name: Jennifer Bryan
  affiliation: |
    RStudio and the Department of Statistics, University of British Columbia
  thanks: |
    The author gratefully acknowledges the constructive feedback from reviewers
    Nicholas Horton, Colin Rundel, and Hadley Wickham.
blinded: 0
output:
  github_document:
    toc: yes
  rticles::asa_article: null
keywords: |
    data science, workflow, Git, GitHub, R language, R Markdown, reproducibility
abstract: |
    Data analysis, statistical research, and teaching statistics have at least
    one thing in common: these activities all produce many files! There may be
    data files, source code, figures, tables, prepared reports, and much more.
    These files generally evolve over the course of a project and often need to
    be shared with others, for reading or edits, as a project unfolds. Without
    explicit and structured management, project organization can easily descend
    into chaos, taking time away from the primary work and reducing the quality
    of the final product. This unhappy result can be avoided by repurposing
    tools and workflows from the software development world, namely, distributed
    version control. This article describes the use of the version control
    system Git and and the hosting site GitHub for statistical and data
    scientific workflows. Special attention is given to projects that use the
    statistical language R and, optionally, R Markdown documents.
bibliography: git-github-for-stats.bib
---

```{r knitr_options, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, out.width = "100%")
is_github_document <-
  knitr::opts_knit$get("rmarkdown.pandoc.to") == "markdown_github"
```

# Why Git?

@@

# What is Git?

@@

# Who should read this and what to expect

@@

# What is GitHub?

@@

# Initial system setup

@@

  * I repeat: consider using a graphical front-end for Git, a.k.a. a Git client, versus restricting yourself to the command line interface.
  * Establish confidence in the basics (e.g. make a change, commit it, push it) before wading into more advanced usage (e.g. branching).
  * Commit yourself to Git usage on a project that will provide sustained practice over several months. Usage in a course is great, because it provides a relentless stream of small deadlines.
  * Realize that no one is giving out Git style points. It's OK to "power-cycle", i.e. re-initialize the Git repository, to get unstuck.

# Repositories and workflow

@@

# Commits, diffs, and tags

@@

# Markdown is special on GitHub

@@

# Markdown is special for R users

@@

Before we move on, I want to zoom out and revisit R Markdown and the rmarkdown package more generally. Note that R Markdown can be rendered to many more formats than Markdown, including HTML, PDF, and Microsoft Word, and can incorporate code chunks in many languages other than R. Markdown is emphasized here because it is extraordinarily useful in GitHub-hosted projects. Many people are used to targeting other output formats and overlook this special synergy between Markdown and GitHub.

# Which files to commit

@@

__Intermediates:__ These files are programmatically generated and serve a temporary purpose, but are not intrinsically valuable (e.g., `.aux` and `.log` in LaTeX workflows).

@@

# Collaboration

@@

# GitHub as web presence

@@

# Where to go next?

I've tried to convey the main points about the use of Git and GitHub in statistical and data analytical settings, but I've left many things out. There are more advanced topics that will come up as your use of Git becomes more sophisticated and there are topics that are only relevant to certain types of reader.

@@

Branches and pull requests are an extremely powerful feature of Git/GitHub and should be your first foray beyond the basics of commit, push, and pull. A branch is an independent line of development within a repo, where the intent is usually to merge it into the master branch when ready. In fact, Figure \ref{fig:diy-vs-git}D depicts exactly such a branch-and-merge workflow, which can be used in both collaborative and solo projects. In a course website, you might work in a branch to update a series of lessons, while leaving the current version intact in the meantime. A pull request is a GitHub-specific way to propose changes to a repo that overlays Git's regular branch and merge workflow with facilities for structured review. A pull request can be made between two branches in the same repo, such as `master` and `bugfix`, or between two independent copies of a repo, an original and a so-called "fork". This is the mechanism for making and accepting contributions to open source software projects on GitHub, including many popular R packages.

# Call to action

Statistics is an important part of Data Science, though there is lively debate about the exact relationship, both past and future [@donoho]. I strongly disagree with glib claims that "Data Science is just Statistics", which I've heard in academia. For reasons both genuine and pragmatic, Statistics departments want to participate in the growth of Data Science at their respective universities. So far, their success has been decidedly mixed [@donoho]. In the era of Data Science, staying current can't be limited to mathematical and methodological developments. The holistic work style described here is widely regarded as current best practice, yet we generally neither practice nor preach it in academic statistics. It is but one example of the many ways in which Data Science is **not** just statistics. By acknowledging the importance of modern workflows and tooling, Statistics can solidify its claim to being one of the core disciplines of Data Science.

```{r eval = is_github_document, echo = FALSE, results = 'asis'}
## surely there is a better way?
cat("# References\n")
```

### 2017-07-07 README.tex
Abstract, conclusion, Header 1 not 2, spellcheck

\bigskip
\begin{abstract}
Abstract abstract abstract.
\end{abstract}

\noindent%
{\it Keywords:} version control
\vfill

@@

\subsection{Why Git?}\label{why-git}

@@

\subsection{What is Git?}\label{what-is-git}

@@

\subsection{Who should read this and what to
expect}\label{who-should-read-this-and-what-to-expect}

@@

\subsection{What is GitHub?}\label{what-is-github}

@@

\subsection{Initial system setup}\label{initial-system-setup}

@@

\begin{itemize}
\tightlist
\item
  I repeat: consider using a graphical front-end for Git, a.k.a. a Git
  client, versus restricting yourself to the command line interface.
\item
  Establish confidence in the basics (e.g.~make a change, commit it,
  push it) before wading into more advanced usage (e.g.~branching).
\item
  Commit yourself to Git usage on a project that will provide sustained
  practice over several months. Usage in a course is great, because it
  provides a relentless stream of small deadlines.
\item
  Realize that no one is giving out Git style points. It's ok to
  ``power-cycle'', i.e.~re-initialize the Git repository, to get
  unstuck.
\end{itemize}

\subsection{Repositories and workflow}\label{repositories-and-workflow}

@@

\subsection{Commits, diffs, and tags}\label{commits-diffs-and-tags}

@@

\subsection{Markdown is special on
GitHub}\label{markdown-is-special-on-github}

@@

\subsection{Markdown is special for R
users}\label{markdown-is-special-for-r-users}

@@

Before we move on, I want to zoom out and revisit R Markdown and the
rmarkdown package more generally. Note that R Markdown can be rendered
to many more formats than Markdown, including HTML, PDF, and Microsoft
Word, and can incorporate code chunks in many languages other than R.
Markdown is emphasized here because it is extraordinarily useful in
GitHub-hosted projects. Many people are used to targetting other output
formats and overlook this special synergy between Markdown and GitHub.

\subsection{Which files to commit}\label{which-files-to-commit}

@@

\textbf{Intermediates:} These files are programmatically generated and
serve a temporary purpose, but are not intrinsically valuable (e.g.,
\texttt{.aux} and \texttt{.log} in LaTex workflows).

@@

\subsection{Collaboration}\label{collaboration}

@@

\subsection{GitHub as web presence}\label{github-as-web-presence}

@@

\subsection{More resources}\label{more-resources}

I've tried to convey the main points about the use of Git and GitHub in
statistical and data analytical settings, but I've had to leave many
things out. There are more advanced topics that will come up once your
use of Git becomes more sophisticated and there are topics that are only
relevant to certain types of reader.

@@

Branches and pull requests are an extremely powerful feature of
Git/GitHub and should be your first foray beyond the basics of commit,
push, and pull. A branch is an independent line of development within a
repo, where the intent is usually to merge it into the master branch
when ready. In a course website, you might work in a branch to update a
series of lessons, while leaving the current version intact in the
meantime. A pull request is a GitHub-specific way to propose changes to
a repo that augments Git's regular branch and merge workflow with
facilities for structured review. A pull request can be made between two
branches in the same repo, such as \texttt{master} and \texttt{bugfix},
or between two independent copies of a repo, an original and a so-called
``fork''. This is the mechanism for making and accepting contributions
to open source software projects on GitHub, including many popular R
packages \emph{link to a search?}.

\emph{inspiration for possible figure on collaboration/mergin, from the
barlett talk}

\includegraphics[width=1\linewidth]{img/bartlett-merge-commit}

\subsection{Conclusion}\label{conclusion}

\emph{NEEDS ONE \ldots{} BUT SHORT \ldots{} SOME FODDER}

Transparency about process and product is increasingly important in
science. The SOMETHING for reproducibility is well accepted. A more
underappreciated benefit is democratization of our field, as this
affords a much broader audience a clear view of how scientists and
programmers work.

make statistical thought and implementation available

>>>>

\bigskip
\begin{abstract}
Data analysis, statistical research, and teaching statistics have at
least one thing in common: these activities all produce many files!
There may be data files, source code, figures, tables, prepared reports,
and much more. These files generally evolve over the course of a project
and often need to be shared with others, for reading or edits, as a
project unfolds. Without explicit and structured management, project
organization can easily descend into chaos, taking time away from the
primary work and reducing the quality of the final product. This unhappy
result can be avoided by repurposing tools and workflows from the
software development world, namely, distributed version control. This
article describes the use of the version control system Git and and the
hosting site GitHub for statistical and data scientific workflows.
Special attention is given to projects that use the statistical language
R and, optionally, R Markdown documents.
\end{abstract}

\noindent%
{\it Keywords:} data science, workflow, Git, GitHub, R language, R Markdown,
reproducibility
\vfill

@@

\section{Why Git?}\label{why-git}

@@

\section{What is Git?}\label{what-is-git}

@@

\section{Who should read this and what to
expect}\label{who-should-read-this-and-what-to-expect}

@@

\section{What is GitHub?}\label{what-is-github}

@@

\section{Initial system setup}\label{initial-system-setup}

@@

\begin{itemize}
\tightlist
\item
  I repeat: consider using a graphical front-end for Git, a.k.a. a Git
  client, versus restricting yourself to the command line interface.
\item
  Establish confidence in the basics (e.g.~make a change, commit it,
  push it) before wading into more advanced usage (e.g.~branching).
\item
  Commit yourself to Git usage on a project that will provide sustained
  practice over several months. Usage in a course is great, because it
  provides a relentless stream of small deadlines.
\item
  Realize that no one is giving out Git style points. It's OK to
  ``power-cycle'', i.e.~re-initialize the Git repository, to get
  unstuck.
\end{itemize}

\section{Repositories and workflow}\label{repositories-and-workflow}

@@

\section{Commits, diffs, and tags}\label{commits-diffs-and-tags}

@@

\section{Markdown is special on
GitHub}\label{markdown-is-special-on-github}

@@

\section{Markdown is special for R
users}\label{markdown-is-special-for-r-users}

@@

Before we move on, I want to zoom out and revisit R Markdown and the
rmarkdown package more generally. Note that R Markdown can be rendered
to many more formats than Markdown, including HTML, PDF, and Microsoft
Word, and can incorporate code chunks in many languages other than R.
Markdown is emphasized here because it is extraordinarily useful in
GitHub-hosted projects. Many people are used to targeting other output
formats and overlook this special synergy between Markdown and GitHub.

\section{Which files to commit}\label{which-files-to-commit}

@@

\textbf{Intermediates:} These files are programmatically generated and
serve a temporary purpose, but are not intrinsically valuable (e.g.,
\texttt{.aux} and \texttt{.log} in LaTeX workflows).

@@

\section{Collaboration}\label{collaboration}

@@

\section{GitHub as web presence}\label{github-as-web-presence}

@@

\section{Where to go next?}\label{where-to-go-next}

I've tried to convey the main points about the use of Git and GitHub in
statistical and data analytical settings, but I've left many things out.
There are more advanced topics that will come up as your use of Git
becomes more sophisticated and there are topics that are only relevant
to certain types of reader.

@@

Branches and pull requests are an extremely powerful feature of
Git/GitHub and should be your first foray beyond the basics of commit,
push, and pull. A branch is an independent line of development within a
repo, where the intent is usually to merge it into the master branch
when ready. In fact, Figure \ref{fig:diy-vs-git}D depicts exactly such a
branch-and-merge workflow, which can be used in both collaborative and
solo projects. In a course website, you might work in a branch to update
a series of lessons, while leaving the current version intact in the
meantime. A pull request is a GitHub-specific way to propose changes to
a repo that overlays Git's regular branch and merge workflow with
facilities for structured review. A pull request can be made between two
branches in the same repo, such as \texttt{master} and \texttt{bugfix},
or between two independent copies of a repo, an original and a so-called
``fork''. This is the mechanism for making and accepting contributions
to open source software projects on GitHub, including many popular R
packages.

\section{Call to action}\label{call-to-action}

Statistics is an important part of Data Science, though there is lively
debate about the exact relationship, both past and future
\citep{donoho}. I strongly disagree with glib claims that ``Data Science
is just Statistics'', which I've heard in academia. For reasons both
genuine and pragmatic, Statistics departments want to participate in the
growth of Data Science at their respective universities. So far, their
success has been decidedly mixed \citep{donoho}. In the era of Data
Science, staying current can't be limited to mathematical and
methodological developments. The holistic work style described here is
widely regarded as current best practice, yet we generally neither
practice nor preach it in academic statistics. It is but one example of
the many ways in which Data Science is \textbf{not} just statistics. By
acknowledging the importance of modern workflows and tooling, Statistics
can solidify its claim to being one of the core disciplines of Data
Science.
